A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-11 docs/icccn2009/paper.tex
This is an intermediate commit: Added more citations to lit.bib and the package "url" to paper.tex. Also added citation references in the document itself.
git-svn-id: svn://localhost/netbeams/trunk@362 559edde2-a4d4-44e0-96fc-4035e40d7396

\usepackage{epsfig}
\usepackage{fancyvrb}

@@

The YSI 6600EDS V2 Sonde is a water quality monitoring device that
gathers water quality data, and operates in fresh, sea, or polluted
water. Some of the measurements that the YSI is capable of are
conductivity, temperature, chloride, ammonium, nitrate, turbidity, and
chlorophyll.  It is under 55 cm in length, 8.9 cm in diameter, and
weighs approximately 3.18 kg. It operates at temperatures between -5
to +50 deg C and at depths up to 200 meters.  It uses 8 C-size
Alkaline Batteries or External 12 VDC. Battery life can last up to 75
days depending on sensor configuration. The Sonde's bulkhead contains
multiple pin connectors to support sensor probes for measuring the
following:

@@

A DSP Component encapsulates a particular functionality. We
distinguish between those components that produce data (DP) and those
that consume data (DC). An example for a DSP Component acting in the
role of a DP is a sensor module. The sensor module reads data from the
physical sensor and converts the data into a message that is forwarded
to the DSP. Because this module generates data with respect to the
DSP, it is a data producer. An example for a data consumer is a
converter module that converts the internal, DSP-specific data format,
suitable for consumption by external applications. One such common
format used in the domain of environmental monitoring is OpenDAP
(REF). Since the converter module accepts messages from the DSP, it
acts in the role of a data consumer. Finally, a DSP Component
implementing persistency is an example of a module that acts in the
roles of both data producer and consumer. It accepts messages from the
DSP and stores them in a database. At a later time, it may act as a
data producer by resending previously stored messages.

@@

Figure \ref{FIG_NETBEAMS} provides an end-to-end overview of the
NetBEAMS architecture. NetBEAMS is built using multiple installation
of the DSP and by providing special purpose components for the various
tasks. One DSP installation is collocated with each sensor and an
additional DSP is running on the backend server. Because of the remote
locations where each individual sensor is deployed in the field and
the sparse nature of the network, the resulting topology is a tree
structure of depth 1.

@@

NetBEAMS supports environmental sensor such as the YSI Sonde
introduced in Section \ref{SEC_BACKGROUND}. In order to access its
sensor readings, we collocate it with an off-the-shelf embedded
computing platform called Gumstix (REF) (far left in Figure
\ref{FIG_NETBEAMS}). The Gumstix is a single-board computers with 256
Mb of flash and 256 Mb of SDRAM, and running a 600 MHz TI OMAP 3503
processor. Additional features can be added to the motherboard with
expansion cards connected via on-board buses. The motherboards draw
less than 250 mA @4V at 400 MHz and less than 50 mA while idling,
waiting for input.  Gumstix runs Linux 2.6 with the BusyBox utilities,
and use the OpenEmbedded build environment to provide a complete Linux
environment and a large range of Linux applications. We use the JamVM
(REF) implementation of Sun Microsystem's virtual machine in order to
run the DSP

We make use of the cellular phone network to establish a communication
link with the backend server. The cellular phone network has the
advantages of reaching well beyond telemetry options such as wireless
networks and being a significantly lower cost solution than satellite
communications. We make use of the Huawei E220. The E220 is a USB
modem that supports HSDPA, UMTS, EDGE packet data services at maximum
transmission rates of 3.6Mbps, 384kbps, and 236.8kbps respectively.
HSDPA and UMTS operate at 2100MHz while GSM, GPRS, and EDGE operate at
900, 1800, and 1900MHz. The modem connects to the Verdex Pro XM4
motherboard via a mini USB interface (supporting USB 2.0). Linux
kernels since 2.6.20 include support for E220 drivers.  This includes
the linux kernel on the Gumstix which runs 2.6.21.  The Gumstix
communicates with the modem using the PPP protocol and accesses the
cellular network using a data plan from AT\&T.

@@

Our implementation of the DSP is based on the Open Source
implementation of OSGi called Knopflerfish (REF). Since Sun
Microsystems does not support its JDK for the Gumstix, we use JamVM
(REF) as the Java virtual machine implementation and GNU Classpath
(REF) for the runtime libraries. We have implemented various DSP
Components that implement the functionality needed for NetBEAMS. On
the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in our
case the aforementioned YSI Sonde. A YSI Sonde data procuder (1) is a DSP
component that uses the RS232 to read the physical measurements from
the device. This component converts the physical measurement into a
DSP-specific message. The following XML Schema defines the YSI Sonde
specific message that captures all measurement points mentioned in
Table \ref{TAB_SONDE_MEASUREMENTS}:

>>>>

\usepackage{epsfig}
\usepackage{fancyvrb}
\usepackage{url}

@@

The YSI 6600EDS V2 Sonde \cite{Sonde01} is a water quality monitoring device that
gathers water quality data, and operates in fresh, sea, or polluted
water. Some of the measurements that the YSI is capable of are
conductivity, temperature, chloride, ammonium, nitrate, turbidity, and
chlorophyll.  It is under 55 cm in length, 8.9 cm in diameter, and
weighs approximately 3.18 kg. It operates at temperatures between -5
to +50 deg C and at depths up to 200 meters.  It uses 8 C-size
Alkaline Batteries or External 12 VDC. Battery life can last up to 75
days depending on sensor configuration. The Sonde's bulkhead contains
multiple pin connectors to support sensor probes for measuring the
following:

@@

A DSP Component encapsulates a particular functionality. We
distinguish between those components that produce data (DP) and those
that consume data (DC). An example for a DSP Component acting in the
role of a DP is a sensor module. The sensor module reads data from the
physical sensor and converts the data into a message that is forwarded
to the DSP. Because this module generates data with respect to the
DSP, it is a data producer. An example for a data consumer is a
converter module that converts the internal, DSP-specific data format,
suitable for consumption by external applications. One such common
format used in the domain of environmental monitoring is OpenDAP
\cite{opendap01}. Since the converter module accepts messages from the DSP, it
acts in the role of a data consumer. Finally, a DSP Component
implementing persistency is an example of a module that acts in the
roles of both data producer and consumer. It accepts messages from the
DSP and stores them in a database. At a later time, it may act as a
data producer by resending previously stored messages.

@@

Figure \ref{FIG_NETBEAMS} provides an end-to-end overview of the
NetBEAMS architecture. NetBEAMS is built using multiple installations
of the DSP and by providing special purpose components for the various
tasks. One DSP installation is collocated with each sensor and an
additional DSP is running on the backend server. Because of the remote
locations where each individual sensor is deployed in the field and
the sparse nature of the network, the resulting topology is a tree
structure of depth 1.

@@

NetBEAMS supports an environmental sensor such as the YSI Sonde,
introduced in Section \ref{SEC_BACKGROUND}. In order to access its
sensor readings, we collocate it with an off-the-shelf embedded
computing platform called Gumstix \cite{gumstix01} (far left in Figure
\ref{FIG_NETBEAMS}). The Gumstix is a single-board computers with 256
Mb of flash and 256 Mb of SDRAM, and running a 600 MHz TI OMAP 3503
processor. Additional features can be added to the motherboard with
expansion cards connected via on-board buses. The motherboards draw
less than 250 mA @4V at 400 MHz and less than 50 mA while idling,
waiting for input.  Gumstix runs Linux 2.6 with the BusyBox utilities,
and uses the OpenEmbedded build environment to provide a complete Linux
environment and a large range of Linux applications. We use the JamVM
\cite{jamvm01} implementation of Sun Microsystem's virtual machine in order to
run the DSP.

The Gumstix connects to the RS232 serial port of the YSI Sonde via via a
null-modem cable.

We make use of the cellular phone network to establish a communication
link with the backend server. The cellular phone network has the
advantages of reaching well beyond telemetry options such as wireless
networks and being a significantly lower cost solution than satellite
communications. We make use of the Huawei E220. The E220 is a USB
modem that supports HSDPA, UMTS, and EDGE packet data services at maximum
transmission rates of 3.6Mbps, 384kbps, and 236.8kbps respectively.
HSDPA and UMTS operate at 2100MHz while GSM, GPRS, and EDGE operate at
900, 1800, and 1900MHz. The modem connects to the Verdex Pro XM4
motherboard via a mini USB interface (supporting USB 2.0). Linux
kernels since 2.6.20 include support for E220 drivers.  This includes
the linux kernel on the Gumstix which runs 2.6.21.  The Gumstix
communicates with the modem using the PPP protocol and accesses the
cellular network using a data plan from AT\&T.

@@

Our implementation of the DSP is based on the Open Source
implementation of OSGi called Knopflerfish \cite{knopflerfish01}. Since Sun
Microsystems does not support its JDK for the Gumstix, we use JamVM
\cite{jamvm01} as the Java virtual machine implementation and GNU Classpath
\cite{classpath01} for the runtime libraries. We have implemented various DSP
Components that implement the functionality needed for NetBEAMS. On
the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in our
case the aforementioned YSI Sonde. A YSI Sonde data producer (1) is a DSP
component that uses the RS232 standard to read the physical measurements from
the device over a serial port. This component converts the physical measurements 
into a DSP-specific message. The following XML Schema defines the YSI Sonde
specific message that captures all measurement points mentioned in
Table \ref{TAB_SONDE_MEASUREMENTS}:
